---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message=FALSE, 
  warning=FALSE
)
```


# zoltr - An R client for the Zoltar data repository API

<!-- badges: start -->
<!-- badges: end -->

This package contains functions for working with the (https://www.zoltardata.com/)[Zoltar] forecast repository's API,
including projects, models, forecasts, and scores. Read more at the [zoltr pkgdown site](todo).


## Installation

You can install the released version of zoltr from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("zoltr")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("reichlab/zoltr")
```


## Usage

Read more at the [zoltr pkgdown site](todo), but briefly you use the `new_connection()` function to create a connection
to (https://www.zoltardata.com/)[Zoltar] and then pass that connection along with the _ID_ of the resource of interest
(e.g., a project, model, or forecast) to this package's various functions like `projects()`, `project_info()`, or
`scores()`. 

```{r example}
library(zoltr)
conn <- new_connection()
conn

the_projects <- projects(conn)
project_id <- the_projects[1,]$id
the_project_info <- project_info(conn, project_id)
names(the_project_info)
the_project_info$name
```

