---
title: "Getting Started"
author: "Matthew Cornell"
date: "2019-05-28"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE, 
  warning=FALSE,
  eval = nzchar(Sys.getenv("IS_DEVELOPMENT_MACHINE"))
)
```


# Getting Started with zoltr

zoltr is an R package that simplifies access to the [zoltardata.com](https://www.zoltardata.com/) API. This vignette 
takes you through the package's main features. So that you can experiment without needing a Zoltar account, we use our 
"flagship" project for the CDC Flu challenge, which should always be available for public read-only access.


## Connect to the host

The starting point for working with Zoltar's API is a `ZoltarConnection` object, obtained via the `new_connection`
function. Most zoltr functions take a `ZoltarConnection` along with the _ID_ of the thing of interest, e.g., a
project, model, or forecast ID. You can obtain the ID using one of the `*_info` functions, and you can always use the
web interface to navigate to the resource and look at its ID in the browser address field. The ID should be self evident
for all resources. For example, the project [CDC Flu challenge](https://www.zoltardata.com/project/4)'s URL is
[https://www.zoltardata.com/project/4](https://www.zoltardata.com/project/4), and you can see that its ID is _4_.

(Note: To access protected resources, you need to call the `zoltar_authenticate()` function before calling any zoltr
functions. But this is unnecessary in this vignette because we're accessing a public project, models, and forecasts.)


<!-- run but don't show. NB: will run only if above eval=TRUE -->
```{r setup, include=FALSE}
library(httr)  # o/w devtools::check() gets `could not find function "POST"` error
```


```{r}
library(zoltr)
conn <- new_connection()
conn
```


## Get a list of all projects on the host

Use the `projects()` function to get a connection's projects as a data.frame. Note that it will only list those that you
are authenticated to access.

```{r}
the_projects <- projects(conn)
str(the_projects)
```


## Get a project to work with, list its info and models, and download its scores

Let's start by getting a public project to work with. We will search the projects list for it by name. Then we will pass
its ID to the `project_info()` function to get details, and pass it to the `models()` function to get a data.frame of
its models.

```{r}
project_id <- the_projects[the_projects$name == "CDC Flu challenge", 'id']  # integer(0) if not found, which is an invalid project ID
the_project_info <- project_info(conn, project_id)
names(the_project_info)
the_project_info$description

the_models <- models(conn, project_id)  # may take some time
str(the_models)

score_data <- scores(conn, project_id)
score_data
```


## Get a model to work with and then list its info and forecasts (if any)

Now let's work with a particular model, getting its ID by name and then passing it to the `model_info()` function to get 
details. Then use the `forecasts()` function to get a data.frame of that model's forecasts.

```{r}
model_id <- the_models[the_models$name == "SARIMA model with seasonal differencing", 'id']  # integer(0) if not found
the_model_info <- model_info(conn, model_id)
names(the_model_info)
the_model_info$description

the_forecasts <- forecasts(conn, model_id)
str(the_forecasts)
```


## Finally, download the first forecast's data in two different formats

You can get forecast data using the `forecast_data()` function, which supports a nested `list` or a `data.table`
(tabular) format.

First as a `list`. (Please see [here](https://www.zoltardata.com/docs#forecast-data-format) for nested format, AKA JSON,
details.) 

```{r}
first_forecast_id <- the_forecasts[1, 'id']  # assumes at least one exists

forecast_data <- forecast_data(conn, first_forecast_id, is_json=TRUE)
forecast_data$locations[[1]]$name
forecast_data$locations[[1]]$targets[[1]]$name
forecast_data$locations[[1]]$targets[[1]]$bins[[1]]
```

And as a `data.frame`:

```{r}
forecast_data <- suppressMessages(forecast_data(conn, first_forecast_id, is_json=FALSE))
forecast_data
```
