---
title: "Getting Started"
author: "Matthew Cornell"
date: "2020-04-01"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE, 
  warning=FALSE,
  eval = nzchar(Sys.getenv("IS_DEVELOPMENT_MACHINE"))
)
```


# Getting Started with zoltr

zoltr is an R package that simplifies access to the [zoltardata.com](https://www.zoltardata.com/) API. This vignette 
takes you through the package's main features. So that you can experiment without needing a Zoltar account, we use the
example project from [docs.zoltardata.com](https://docs.zoltardata.com/), which should always be available for public
read-only access.

**NOTE: You will need an account to access the zoltar API - please see [docs.zoltardata.com](https://docs.zoltardata.com/)
for details.**


## Connect to the host and authenticate

The starting point for working with Zoltar's API is a `ZoltarConnection` object, obtained via the `new_connection`
function. Most zoltr functions take a `ZoltarConnection` along with the API _URL_ of the thing of interest, e.g., a
project, model, or forecast. API URLs look like `https://www.zoltardata.com/api/project/3/`, which is that of the
[Docs Example Project](https://www.zoltardata.com/project/3). An important note regarding URLs:

    zoltr's convention for URLs is to require a trailing slash character ('/') on all URLs. The only exception is the
    optional `host` parameter passed to `new_connection()`. Thus, `https://www.zoltardata.com/api/project/3/` is valid,
    but `https://www.zoltardata.com/api/project/3` is not.

You can obtain a URL using some of the `*_info` functions, and you can always use the web interface to navigate to the
item of interest and look at its URL in the browser address field. Keep in mind that you'll need to add `api` to the
browsable address, along with the trailing slash character. For example, you'd browse the _Docs Example Project_ project
at `https://www.zoltardata.com/project/3`, but it's API for use in zoltr would be
`https://www.zoltardata.com/api/project/3/`.

As noted above, all API calls require an account. To access your project, you'll first need to authenticate via the
`zoltar_authenticate()` function. Pass it the username and password for your account. Notes:

- Be careful to store and use your username and password so that they're not accessible to others. A good practice is to
  put them them in your `.Renviron` file and then use `Sys.getenv()` to retrieve them, as outlined in
  [Use Environment variables](https://db.rstudio.com/best-practices/managing-credentials/#use-environment-variables) and
  as done below.
- The Zoltar service uses a "token"-based scheme for authentication. These tokens have a five minute expiration for
  security, which requires re-authentication after that period of time. The zoltr library takes care of 
  re-authenticating as needed by passing your username and password back to the server to get another token. Note that
  the connection object returned by the `new_connection` function stores a token internally, so be careful if saving
  that object into a file.
- Below you will see the host shown as _example.com_ - this is a temporary one that was used to generate this 
  documentation.

For this and other vignettes, you will need to create an [`.Renviron`](https://csgillespie.github.io/efficientR/3-3-r-startup.html#renviron)
file that contains `Z_USERNAME` and `Z_PASSWORD` variables that match your account settings (note the `Z_` prefix).
Then you'll be able to create an authenticated connection:

<!-- run but don't show. NB: will run only if above eval=TRUE -->
```{r setup, include=FALSE}
library(httr)  # o/w devtools::check() gets `could not find function "POST"` error
```

<!-- run but don't show. NB: will run only if above eval=TRUE -->
```{r, include=FALSE}
library(zoltr)
zoltar_connection <- new_connection(host = Sys.getenv("Z_HOST"))
zoltar_authenticate(zoltar_connection, Sys.getenv("Z_USERNAME"), Sys.getenv("Z_PASSWORD"))
zoltar_connection
```

<!-- show but don't run-->
```{r, eval=FALSE, include=TRUE}
library(zoltr)
zoltar_connection <- new_connection()
zoltar_authenticate(zoltar_connection, Sys.getenv("Z_USERNAME"), Sys.getenv("Z_PASSWORD"))
zoltar_connection
```


## Get a list of all projects on the host

Now that you have a connection, you can use the `projects()` function to get all projects as a `data.frame`. Note that
it will only list those that you are authorized to access, i.e., all public projects plus any private ones that you
own or are a model owner.

```{r}
the_projects <- projects(zoltar_connection)
str(the_projects)
```


## Get a project to work with and list its info, models, and scores

Let's start by getting a public project to work with. We will search the projects list for it by name. Then we will pass
its URL to the `project_info()` function to get a `list` of details, and then pass it to the `models()` function to get
a `data.frame` of its models.

```{r}
project_url <- the_projects[the_projects$name == "Docs Example Project", "url"]
the_project_info <- project_info(zoltar_connection, project_url)
names(the_project_info)
the_project_info$description

the_models <- models(zoltar_connection, project_url)
str(the_models)
```

There is other project-related information that you can access, such as its configuration (`zoltar_units()`, 
`targets()`, and `timezeros()` - concepts that are explained at [docs.zoltardata.com](https://docs.zoltardata.com/) ),
`scores()` and `truth()`. As an example, let's get its score data. (Note that available scores are limited due to the
nature of the example project.)

```{r}
score_data <- scores(zoltar_connection, project_url)
score_data
```


## Get a model to work with and list its info and forecasts

Now let's work with a particular model, getting its URL by name and then passing it to the `model_info()` function to
get details. Then use the `forecasts()` function to get a `data.frame` of that model's forecasts (there is only one).
Note that obtaining the model's URL is straightforward because it is provided in the `url` column of `the_models`.

```{r}
model_url <- the_models[the_models$name == "docs forecast model", "url"]
the_model_info <- model_info(zoltar_connection, model_url)
names(the_model_info)
the_model_info$name

the_forecasts <- forecasts(zoltar_connection, model_url)
str(the_forecasts)
```


## Finally, download the forecast's data

You can get forecast data using the `download_forecast()` function, which is in a nested `list` format. Please see
[docs.zoltardata.com](https://docs.zoltardata.com/) for forecast format details.

```{r}
forecast_url <- the_forecasts[1, "url"]
forecast_info <- forecast_info(zoltar_connection, forecast_url)
forecast_data <- download_forecast(zoltar_connection, forecast_url)
length(forecast_data$predictions)
```
