---
title: "Project Owners"
author: "Matthew Cornell"
date: "2019-05-16"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Project Owners}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE, 
  warning=FALSE,
  eval = nzchar(Sys.getenv("IS_DEVELOPMENT_MACHINE"))
)
```


# Zoltr for Project Owners
Welcome to the zoltr vignette for project owners and forecasters. You should read this if you are interested in creating
and managing your own [zoltardata.com](https://www.zoltardata.com/) projects using this package to access them via the 
Zoltar API. Building on the _Getting Started_ vignette, this one covers deleting and uploading forecasts. Note that
before starting you should:

- Create an account on [zoltardata.com](https://www.zoltardata.com/).
- Create a 'sandbox' project with one timezero for testing.
- Create a model in that project.
- Edit your [`.Renviron`](https://csgillespie.github.io/efficientR/3-3-r-startup.html#renviron) file:
    - `USERNAME` and `PASSWORD`: should match your account settings
    - `PROJECT_NAME` and `MODEL_NAME`: should match your project and model
    - `TIMEZERO_DATE`: should match your project
    - `FORECAST_CSV_FILE`: the path of a forecast data file in the format documented in
      [Zoltar documentation](https://www.zoltardata.com/docs)


## Connect to the host and authenticate

To access your project, you'll first need to authenticate via the `zoltar_authenticate()` function. Pass it the username
and password for your account. Note that currently Zoltar uses a five minute timeout on the underlying JWT tokens used
under the hood, which means you'll have to re-authenticate after that time.

(Note: Below you will see the host shown as _127.0.0.1:8000_ - this is a temporary one that was used to run this 
documentation against.)


<!-- run but don't show. NB: will run only if above eval=TRUE -->
```{r setup, include=FALSE}
library(httr)  # o/w devtools::check() gets `could not find function "POST"` error
```


<!-- run but don't show. NB: will run only if above eval=TRUE -->
```{r, include=FALSE}
library(zoltr)
conn <- new_connection(host = Sys.getenv("HOST"))
zoltar_authenticate(conn, Sys.getenv("USERNAME"), Sys.getenv("PASSWORD"))
conn
```

<!-- show but don't run-->
```{r, eval=FALSE, include=TRUE}
library(zoltr)
conn <- new_connection()
zoltar_authenticate(conn, Sys.getenv("USERNAME"), Sys.getenv("PASSWORD"))
conn
```


## Get the sandbox project and model work with and then list its forecasts (if any)

Get the ID of the project named by PROJECT_NAME and then get the ID of that project's model named MODEL_NAME. From the
model ID, call the `forecasts()` function to get a data.frame of that model's forecasts.

```{r}
the_projects <- projects(conn)
project_id <- the_projects[the_projects$name == Sys.getenv("PROJECT_NAME"), 'id']  # integer(0) if not found, which is an invalid project ID
the_models <- models(conn, project_id)
model_id <- the_models[the_models$name == Sys.getenv("MODEL_NAME"), 'id']  # integer(0) if not found
the_forecasts <- forecasts(conn, model_id)
str(the_forecasts)
```


## Delete the model's existing forecast for TIMEZERO_DATE, if any, first printing its info

Get the forecast for the timezero date specified by TIMEZERO_DATE and then call `delete_forecast()`, if one was found.

```{r}
timezero_date_str <- Sys.getenv("TIMEZERO_DATE")
timezero_date <- as.Date(timezero_date_str, format = "%Y%m%d")  # YYYYMMDD
existing_forecast_id <- the_forecasts[the_forecasts$timezero_date == timezero_date, 'id']

the_forecast_info <- forecast_info(conn, existing_forecast_id)  # `Not Found (HTTP 404)` this if forecast doesn't exist 
str(the_forecast_info)

delete_forecast(conn, the_forecast_info$id)
```


## Upload a forecast

Now let's upload the forecast data file FORECAST_CSV_FILE for TIMEZERO_DATE via the `upload_forecast()` function.

Keep in mind that Zoltar queues long operations like forecast uploading, which keeps the site responsive, but makes the
Zoltar API a little more complicated. Rather than having the `upload_forecast()` function _block_ until the upload is 
done, you instead get a quick response in the form of an `UploadFileJob` ID that you can pass to the `upload_info()`
function to check its status to find out when the upload is done (or failed). (This is called _polling_ the host to ask
the status.) Here we poll every second using a helper function.

```{r}
busy_poll_upload_file_job <- function(upload_file_job_id) {
    cat(paste0("polling for status change. upload_file_job: ", upload_file_job_id, "\n"))
    while (TRUE) {
        status <- upload_info(conn, upload_file_job_id)$status
        cat(paste0(status, "\n"))
        if (status == "FAILED") {
            cat(paste0("x failed\n"))
            break
        }
        if (status == "SUCCESS") {
            break
        }
        Sys.sleep(1)
    }
}
```


```{r}
forecast_csv_file <- Sys.getenv("FORECAST_CSV_FILE")
upload_file_job_id <- upload_forecast(conn, model_id, timezero_date_str, forecast_csv_file)
busy_poll_upload_file_job(upload_file_job_id)
```


And here's the upload's final information:

```{r}
the_upload_info <- upload_info(conn, upload_file_job_id)
str(the_upload_info)
```


## Print information about the newly-uploaded forecast

Hopefully you'll see some number of "QUEUED" entries followed by a "SUCCESS" one. Now let's get the new forecast's
information.

Due to there being multiple kinds of files you might upload to Zoltar (including project truth, project template, and
model forecasts), the API returns information about uploaded files in the `$output_json` field of successful upload
job's info. For newly-uploaded forecasts, that field contains a `$forecast_pk` field.

```{r}

new_forecast_id <- upload_info(conn, upload_file_job_id)$output_json$forecast_pk
the_forecast_info <- forecast_info(conn, new_forecast_id)
str(the_forecast_info)
```


## Look at the model's forecast list to see the new forecast

Finally, let's again examine the model's forecasts and notice the new one is there.

```{r}
the_forecasts <- forecasts(conn, model_id)
str(the_forecasts)
